---
title: "Oncogenetic network estimation with Bayesian Networks (OncoBN)"
author: "R Package version 0.1.0."
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an R package that implements methods for inferring the temporal order of somatic mutations in cancer.

## Input data

Input data should be somatic mutation data with samples (or patients) in rows and mutations in the columns. The package includes an example dataset:

```{r}
library(OncoBN)
data("example")
head(example)
```

## Inferring the order of mutations

Use the function `fitCPN()` to fit a Bayesian network to the data. The two model options are conjunctive Bayesian networks (CBN) or disjunctive Bayesian networks (DBN). CBNs assume conjunctive interactions between parent and child nodes (i.e., all parent aberrations must be present in order for the child aberration to occur). DBNs assume disjunctive interactions between parent and child nodes (i.e., at least one of the parent aberrations must be present in order for the child mutation to occur). We can fit both models on the example data:

```{r}
out.cbn <- fitCPN(example,
                  model="CBN",
                  epsilon=0.01)
```

```{r}
out.dbn <- fitCPN(example,
                  model="DBN",
                  epsilon=0.01)
```

The `epsilon` parameter determines the penalty applied to samples that deviate from the estimated network. We recommend choosing `epsilon` to be at most half the prevalence of the least common mutation in the dataset.

Now we can use `plotCPN()` to see the estimated network:

```{r}
plotCPN(out.cbn)
```

```{r}
plotCPN(out.dbn)
```

## Quantifying uncertainty

`fitCPN()` returns a point estimate of the true progression network. To quantify uncertainty in this network we can perform resampling to obtain bootstrap estimates for the network. This is provided by `bootstrapCPN()`:

```{r}
bs <- bootstrapCPN(example,model="DBN")

df <- as.data.frame(cbind(bs$Edges, bs$counts/bs$N))
df[,3] <- as.numeric(df[,3])
df
```

In this example we are 100% in each inferred edge. This is usually not the case in real data. One can also plot a graph that overlays high confidence edges, see `plotBootstrapCPN()`. 

## References

For a description of the DBN model and corresponding inference algorithms, see: 

Nicol, P.B., Coombes, K.R., Deaver, C., Chkrebtii, O.A., Paul, S., Toland, A.E., and Asiaee, A. (2021). Oncogenetic network estimation with disjunctive Bayesian networks. *Computational and Systems Oncology* Vol 1(2), pg.\ 1-e1027. [https://doi.org/10.1002/cso2.1027](https://doi.org/10.1002/cso2.1027).




